The Perils of Exploration under Competition: \\ A Computational Modeling Approach

### 2019-12-10 journal paper/content/sec-intro.tex
tweaks to paper

Given the prevalence of these technologies in the digital economy, our main high-level question is: {\bf does competition lead to the adoption of better exploration algorithms?} This translates into a number of more concrete questions. While it is commonly assumed that ``better" technology always helps, is this so for our setting? Does increasing competition lead to higher consumer welfare? To what extent is there a ``data feedback loop" where one firm having more data leads to that firm attracting more users which leads to that firm having more data, etc.?

>>>>

The main high-level question that we focus on in this paper is: {\bf does competition lead to the adoption of better exploration algorithms?} This translates into a number of more concrete questions. While it is commonly assumed that ``better" technology always helps, is this so for our setting? Does increasing competition lead to higher consumer welfare? To what extent is there a ``data feedback loop" where one firm having more data leads to that firm attracting more users which leads to that firm having more data, etc.?

### 2019-12-10 journal paper/content/sec-model.tex
fixup
tweaks to paper

Each principal commits to a learning algorithm for making recommendations. This algorithm follows a protocol of \emph{multi-armed bandits} (\emph{MAB}). Namely, the algorithm proceeds in time-steps:%
\footnote{These time-steps will sometimes be referred to as \emph{local steps/rounds}, so as to distinguish them from ``global rounds" defined before. We will omit the local vs. local distinction when clear from the context.} each time it is called, it outputs a chosen action $a\in A$ and then inputs the reward for this action. The algorithm is called only in global rounds when the corresponding principal is chosen.

@@

We will emphasize several specific algorithms or classes thereof. In the second half of the paper we will :

\begin{itemize}
\item ``smart" MAB algorithms that combine exploration and exploitation, such as UCB1 \cite{bandits-ucb1} and \gaedit{Thompson Sampling \cite{russo2018tutorial}}. These algorithms achieve
        $\BIR(n) \leq  \tilde{O}(n^{-1/2})$
    for all priors and all (or all but a very few) steps $n$. This bound is known to be tight for any fixed $n$.
    \footnote{This follows from the lower-bound analysis in \cite{bandits-exp3}.}

@@

Write
    $\PMR(a\mid E) = \E[\mu_a \mid E]$
for posterior mean reward of action $a$ given event $E$.

>>>>

Each principal commits to a learning algorithm for making recommendations. This algorithm follows a protocol of \emph{multi-armed bandits} (\emph{MAB}). Namely, the algorithm proceeds in time-steps:%
\footnote{These time-steps will sometimes be referred to as \emph{local steps/rounds}, so as to distinguish them from ``global rounds" defined before. We will omit the global vs. local distinction when clear from the context.} each time it is called, it outputs a chosen action $a\in A$ and then inputs the reward for this action. The algorithm is called only in global rounds when the corresponding principal is chosen.

@@

We will emphasize several specific algorithms or classes thereof:

\begin{itemize}
\item ``smart" MAB algorithms that combine exploration and exploitation, such as UCB1 \cite{bandits-ucb1} and \gaedit{Thompson Sampling \cite{russo2018tutorial}}. These algorithms achieve
        $\BIR(n) \leq  \tilde{O}(n^{-1/2})$
    for all priors and all (or all but a very few) steps $n$. This bound is known to be tight for any fixed $n$.\footnote{This follows from the lower-bound analysis in \cite{bandits-exp3}.}

@@

We denote
    $\PMR(a\mid E) = \E[\mu_a \mid E]$
for posterior mean reward of action $a$ given event $E$.

### 2019-12-10 journal paper/ec19paper/content/non_greedy_choice.tex
the latex compiles
tweaks to paper

Let us consider an extension in which the agents' choice is no longer deterministic. Recall that in our main model agents deterministically choose the firm with the higher reputation score; call this choice rule \emph{HardMax} (\HM). Now, we introduce some randomness: each agent selects between the firms uniformly with probability $\eps\in (0,1)$, and takes the firm with the higher reputation score with the remaining probability; call this choice rule \emph{HardMax with randomness} (\HMR).

@@

\begin{finding}\label{find:non_greedy_choice}
\textit{\TS is weakly dominant under the \HMR~ choice rule, if and only if $\eps T$ is sufficiently large. \asedit{\HMR leads to lower variance in market shares, compared to \HM.}}
\end{finding}

>>>>

Let us consider an extension in which the agents' choice is no longer deterministic. Recall that in the reputation choice variant agents deterministically choose the firm with the higher reputation score, which, as before, we denote as \emph{HardMax} (\HM). Now, we introduce some randomness: each agent selects between the firms uniformly with probability $\eps\in (0,1)$, and takes the firm with the higher reputation score with the remaining probability; as before, we denote this rule as \emph{HardMax with randomness} (\HMR).

@@

\begin{finding}\label{find:non_greedy_choice}
\textit{\TS is weakly dominant under the \HMR~ choice rule, if and only if $\eps T$ is sufficiently large. \HMR leads to lower variance in market shares, compared to \HM.}
\end{finding}
